---
title: "Gapminder"
author: "Leo Lu"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---



```{r, message=FALSE}
# Standard data manipulation and visulisation
library(dplyr)
library(ggplot2)

# Tools for working with models
# devtools::install_github("hadley/modelr")
library(modelr)

# Tools for working with lots of models
library(purrr)
library(tidyr)

# Data
library(gapminder)
```

## Data

We can take a quick peak at the loaded `gapminder` object:

This package provides a subset of the data provided by

[Gapminder](http://www.gapminder.org/data/). The data comes as a data frame
(or `tbl_df` if you're using `dplyr`). We can visit [Gapminder.org
documentation page](http://www.gapminder.org/data/documentation/) to obtain
information about the columns:

| variable  | meaning                  |
|:----------|:-------------------------|
| country   |                          |
| continent |                          |
| year      |                          |
| lifeExp   | life expectancy at birth |
| pop       | total population         |
| gdpPercap | per-capita GDP           |

```{r}
library(gapminder)
gapminder
```

## Plot Data

```{r}
gapminder %>% 
  ggplot(aes(year, lifeExp, group = country)) +
    geom_line(alpha = 1/3)
```


## Plot Models

```{r}
nz <- filter(gapminder, country == "New Zealand")
nz %>% 
  ggplot(aes(year, lifeExp)) + 
  geom_line() + 
  ggtitle("Full data = ")

nz_mod <- lm(lifeExp ~ year, data = nz)
nz %>% 
  add_predictions(nz_mod) %>%
  ggplot(aes(year, pred)) + 
  geom_line() + 
  ggtitle("Linear trend + ")

nz %>% 
  add_residuals(nz_mod) %>% 
  ggplot(aes(year, resid)) + 
  geom_hline(yintercept = 0, colour = "white", size = 3) + 
  geom_line() + 
  ggtitle("Remaining pattern")
```

## Refernce

- [Hadley: r4ds](http://r4ds.had.co.nz/many-models.html)
