---
title: "Color palette with ggplot and RColorBrewer"
author: "Leo Lu"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---


```{r setup, include=FALSE}
# set root dir when rendering
knitr::opts_knit$set(root.dir = '..')
knitr::opts_chunk$set(comment = "#>",
                      fig.align='center',
                      collapse = TRUE)
```

本範例需要的套件

```{r, message=FALSE}
library(magrittr)
library(dplyr)
library(ggplot2)
library(RColorBrewer)
```

## 玩玩顏色

```{r}
library(ggplot2)
library(RColorBrewer)

data(mpg)
mpg

p <- ggplot(mpg, aes(x = displ, y = hwy, color = class)) +
  geom_point()
plot(p)
```


我們可以很簡單地用 `scale_color_brewer()` 調整(離散型變數)的填滿顏色

```{r}
p + scale_color_brewer() # default color scheme
```

## Types of palettes

看看有那些(離散變數)色彩主題可挑選

```{r, fig.height=8}
library(RColorBrewer)
display.brewer.all()
```


- Diverging

BrBG, PiYG, PRGn, PuOr, RdBu, RdGy, RdYlBu, RdYlGn, Spectral

- Qualitative

Accent, Dark2, Paired, Pastel1, Pastel2, Set1, Set2, Set3

- Sequential

Blues, BuGn, BuPu, GnBu, Greens, Greys, Oranges, OrRd, PuBu, PuBuGn, PuRd, Purples, RdPu, Reds, YlGn, YlGnBu, YlOrBr, YlOrRd


換成不同顏色主題試試 `scale_color_brewer palette`:

```{r}
p + scale_color_brewer(palette="Set1")
```


## 顏色不夠怎麼辦

```{r}
p2 <- ggplot(mpg) + 
  geom_bar(aes(x = factor(displ), fill = factor(displ)))
p2 + scale_fill_brewer(palette="Set2")
```


Indeed `length(unique(mpg$displ))` finds `r length(unique(mpg$displ))` unique values for horse power in mpg, while specified palette Set2 has 8 colors to choose from. Lack of colors in the palette triggers ggplot warnings like this (and invalidates plot as seen above):

```
1: In brewer.pal(n, pal) :
  n too large, allowed maximum for palette Set2 is 8
Returning the palette you asked for with that many colors
```


## Produce larger palettes: Use `colorRamp {grDevices}`

`RColorBrewer` gives us a way to produce larger palettes by interpolating existing ones with constructor function `colorRampPalette`. It generates functions that do actual job: they build palettes with arbitrary number of colors by interpolating existing palette. To interpolate palette Set1 to 22 colors (number of colors is stored in `colour_count` variable for examples to follow):

```{r}
colour_count = length(unique(mpg$displ))
getPalette = colorRampPalette(brewer.pal(9, "Set1"))
getPalette

p2 + 
  scale_fill_manual(values = getPalette(colour_count)) + 
  theme(legend.position="right")
```


## Legend problem

Another likely problem with large number of bars in histogram plots is placing of the legend. Adjust legend position and layout using theme and guides functions as follows :

```{r}
p2 + 
  scale_fill_manual(values = getPalette(colour_count)) +
  theme(legend.position = "bottom") +
  guides(fill = guide_legend(nrow=2))
```


Finally, the same example using in place palette constructor with different choice of library palette:

```{r}
p2 + 
  scale_fill_manual(values = colorRampPalette(brewer.pal(12, "Accent"))(colour_count)) +
  theme(legend.position="bottom") +
  guides(fill=guide_legend(nrow=2))
```


There are many more [scale functions](http://docs.ggplot2.org/current/scale_brewer.html) to choose from depending on aesthetics type (`colour`, `fill`), color types (`gradient`, `hue`, etc.), data values (`discrete` or `continuous`).

## Refernce

- http://novyden.blogspot.tw/2013/09/how-to-expand-color-palette-with-ggplot.html
